
---
title: "Decision Trees in R"
output: word_document
---

In the CIND119 - Introduction to Big Data course, decision trees were discussed in the week 4 learning materials. Here is an example of decision trees in R.

```{r setup, include=FALSE}
library("party")
```

We will be using the iris data set for this example.
```{r}
head(iris)
str(iris)
```

We will build a decision tree for the iris data set. We will be using the 'ctree' function. The first parameter will be the formula we want to use. Here we will choose 'Species' as the dependent variable and 'Sepal.Length', 'Sepal.Width', 'Petal.Length' and 'Petal.Width' will be the independent variables.
```{r}
iris_ctree <- ctree(Species ~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width, data=iris)
```

Let's print the decision tree.
```{r}
print(iris_ctree)
```

Now, let's plot the decision tree.
```{r}
plot(iris_ctree, type="simple")
```

As we mentioned in the video lectures of module 10, decision trees are also used as classifiers. To this aim, let's split the iris dataset into training and test set.
```{r}
train_index <- sample(1:nrow(iris), 0.7 * nrow(iris))
train.set <- iris[train_index,]
test.set  <- iris[-train_index,]
```

Let's run our model on the training set.
```{r}
iris_ctree_model <- ctree(Species ~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width, data=train.set)
iris_ctree_model
```

Now, let's make our prediction on the test set.
```{r}
iris_ctree_prediction <- predict(iris_ctree_model, test.set) 
# gives the probability for each class
head(iris_ctree_prediction)
```

Let's see the confusion matrix.
```{r}
table(iris_ctree_prediction, test.set$Species)
```